Compilation and Backend-Independent Vectorization for Multi-Party Computation

### 2022-05-01 overleaf/related.tex
Minor edits in these secs.
A comment in Related work.
Added comments.

Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
restrict flow between secure and insecure parts of the program.
%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we
%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our
%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations
%thrust of the proposal.

@@

Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. Buscher~\cite{Buscher:2018} considers SIMD-vectorization
at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.

>>>>

Other works, e.g., Obliv-C~\cite{Zahur:2015}, Wystiria~\cite{SP:RasHamHic14} and Viaduct~\cite{Acay:2021} focus on higher-level language design, particularly information-flow systems that
restrict flow between secure and insecure parts of the program.

%\ana{Add discussion on HyCC/Buscher here. 1. Formalization, reasoning about correctness of transformation. Source-to-source cannot handle if-then-else vs. MPC Source.}
%Key ideas such as ${\sf insecure} <: {\sf secure}$ qualifier systems are similar to the ideas we
%present in~\Secref{sec:interprocedural} and~\Secref{sec:zkp} and we intend to draw on these work. While we will design a high-level language expanding our
%intraprocedural IMP-MPC syntax and semantics into the \emph{interprocedural} case, our interest and focus lie primarily in the ``machine-independent'' optimizations
%thrust of the proposal.

HyCC~\cite{CCS:BDKKS18} is a compiler from C Source into ABY/MOTION circuits. It does source-to-source compilation with the key goal to decompose the 
program into modules and then assign protocols to modules. In contrast, we focus on MPC Source-level optimizations, specifically vectorization, although we envision 
future optimizations as well. We formalize MPC Source and reasoning about transformations, which we conjecture is more tractable than reasoning over
the higher-level AST. On the other hand, HyCC does inter-procedural optimizations, while our analysis is intra-procedural. We will explore context-sensitive inter-procedural 
analysis over MPC Source in future work. HyCC, similarly to Buscher~\cite{Buscher:2018} uses an of-the-shelf source-to-source polyhedral compiler~\footnote{We believe HyCC uses 
Par4All~\ana{add citation!}, however, it is not appear to be included with the publicly available distribution of HyCC.}
to perform vectorization at the level of source code. The disadvantage of using an of-the-shelf source-to-source compiler is that it solves a more general 
problem than what MPC presents and may forgo opportunities for optimization --- specifically, it is well-known that vectorization and polyhedral compilation 
does not work well with conditionals~\cite{Benabderrahmane:2010,Karrenberg:2015}. In contrast, we consider vectorization at the level of MPC Source which 
linearizes conditionals; we are able to handle programs with interleaved if-statements and for-loops and achieve significant speedup.

@@

Furthermore, extensions of classical loop vectorization with array writes, arbitrary indexing, including non-affine indexing, and interaction with SSA are non-trivial
and present novel challenges and opportunities for contribution. Polyhedral parallelization~\cite{Benabderrahmane:2010} considers a higher-level source (typically AST)
representation, while our work takes advantage of linear MPC Source and SSA form. The work by Karrenberg~\cite{Karrenberg:2015} is rare in that space, in the sense that it considers
vectorization over SSA form, which has similarities to MPC Source. We consider different array representation, notion of dependence, and reasoning
about dependence, which we conjecture is more suitable for MPC Source. 
%Buscher~\cite{Buscher:2018} considers SIMD-vectorization
%at the level of source code, which then combines with circuit-level optimizations in the TinyGarble compiler.
%He proposes using an off-the-shelf polyhedral compiler, however, application is limited to only two routines,
%essentially just inner product and euclidian distance; it is unclear how effective the off-the-shelf compiler is.
%In contrast, we consider vectorization at the level of MPC Source separating ``backend-independent'' vectorization
%and circuit-level amortization (done by MOTION). We apply our compiler on a wide range of routines.
